---
title: "R Markdown"
author: "Jiri J. Sebek"
date: "2025-03-09"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{fvextra} % Required for advanced verbatim environments
  - \usepackage{fancyvrb} % Enhances verbatim text (code chunks)
  - \usepackage{upquote} % Ensures correct quote formatting in code
  - \usepackage{xcolor} % Required for syntax highlighting
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
  - \RecustomVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment

## Full Data Description

<http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones>

## Assignment Data

<https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip>

## Assignment

1.  You should create one R script called run_analysis.R that does the following.
2.  Merges the training and the test sets to create one data set.
3.  Extracts only the measurements on the mean and standard deviation for each measurement.
4.  Uses descriptive activity names to name the activities in the data set.
5.  Appropriately labels the data set with descriptive variable names.
6.  From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

### Download data set, unzip, root directory
```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(url, destfile = "dataset.zip")
unzip("dataset.zip")
```

### Load to R
```{r}

```
